---
title: "PA1_template"
date: "18/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part1: Loading and preprocessing the data
```{r Load the data}
data <- read.csv("activity.csv")
head(data)
str(data)
sum(is.na(data$steps))
mean(is.na(data$steps))
```

```{r Process/transform the data}
data$date <- as.Date(as.character(data$date, "%Y-%m-%d"))
data$weekday <- as.factor(weekdays(data$date, abbreviate = FALSE))
str(data)
summary(data)
```


# Q1: What is mean total number of steps taken per day?

## Calculate the total number of steps taken per day
```{r}
totalstep <- aggregate(steps ~ date, data, sum, na.rm=TRUE)
summary(totalstep)
head(totalstep)
```

## Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(totalstep,
       aes(y=totalstep$steps, 
           x=totalstep$date))+
  geom_bar(stat="identity") +
  ylab("Total steps each day")+
  xlab("Date")+
  ggtitle("Total number of steps taken each day")
```

## Calculate and report the mean and median of the total number of steps taken per day

## The mean and median of the total number of steps taken per day are 10766.19 and 10765 respectively. 

```{r overall mean steps / media steps}
overall.mean.step <- mean(totalstep$steps)
overall.mean.step

overall.median.step <- median(totalstep$steps)
overall.median.step
```

```{r mean steps each day}
library(dplyr)
library(tidyr)
mean.step <- data %>% 
  select(date,steps) %>% 
  group_by(date) %>% 
  summarise(mean(steps))%>%
  drop_na()

ggplot(mean.step,
       aes(y=mean.step$`mean(steps)`, 
           x=mean.step$date))+
  geom_bar(stat="identity") +
  ylab("Average steps each day")+
  xlab("Date")+
  ggtitle("Average number of steps taken each day")
```
```{r media steps each day}
media.step <- data %>% 
  select(date,steps) %>% 
  group_by(date) %>% 
  summarise(median(steps))%>%
  drop_na()
```

# Q2: What is the average daily activity pattern?

## Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
intervalstep <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
names(intervalstep) <- c("interval", "averagesteps")
plot(averagesteps ~ interval, 
     data = intervalstep, 
     type = "l",
     xlim=c(0,2500))

```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## The maximum number of steps is in the No.835 5-minute interval

```{r maximum number of steps in 5-minute interval}
intervalstep[which.max(intervalstep$averagesteps), ]$interval
```

# Part2: Imputing missing values

## The presence of missing days may introduce bias into some calculations or summaries of the data.

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

## There are 2304 missing values in the dataset, accounting for about 13.11% of the dataset.  
```{r}
sum(is.na(data$steps))
mean(is.na(data$steps))
```

## Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
fill.values <- mean.step$`mean(steps)`[match(data$interval, mean.step$interval)]

## Create a new dataset that is equal to the original dataset but with the missing data filled in.

fill.data <- transform(data, steps = ifelse(is.na(data$steps), yes = fill.values, no = data$steps))

fill.totalstep <- aggregate(steps ~ date, fill.data, sum)
head(fill.totalstep)

```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

### Do these values differ from the estimates from the first part of the assignment? 
```{r}
## 1st plot with NA data
  ggplot(totalstep,
         aes(y=totalstep$steps,
             x=totalstep$date))+
    geom_bar(stat="identity") +
    ylab("Total steps each day")+
    xlab("Date")+
    ggtitle("Total number of steps taken each day")

## 2nd plot without NA data
  ggplot(fill.totalstep, 
         aes(y=fill.totalstep$steps, 
             x=fill.totalstep$date))+
    geom_bar(stat="identity") +
    ylab("Total steps each day")+
    xlab("Date")+
    ggtitle("Total number of steps taken each day (without NA data)")

```

### What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r compare the mean}
# It seems that there is no difference in terms of the mean. 
mean(totalstep$steps)
mean(fill.totalstep$steps)
```
```{r compare the media}
# It seems that there is no difference in terms of the median.
median(totalstep$steps)
median(fill.totalstep$steps)
```

# Part3: Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
fill.data$weekday <- as.factor(weekdays(fill.data$date, abbreviate = FALSE))
table(fill.data$weekday)

fill.data$DayType <- ifelse(fill.data$weekday=='Saturday' | fill.data$weekday=='Sunday', 'weekend','weekday')
head(fill.data)
```

## Make a panel plot containing a time series plot (i.e.type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
## There are more steps in weekends than in weekdays. 
```{r}
week.step <- aggregate(steps~interval + DayType,
                       data=fill.data,
                       FUN=mean,
                       na.rm=TRUE)

library(lattice)
xyplot(steps ~ interval | DayType, week.step, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```


